An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}

### 2022-09-26 paper/paper1/ms.tex
The biggest thorn in the cloning--- line core residuals

We retain a relatively small number of pixels $N_{\rm cut}$ adjacent to the line center. Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra. The choice of $N_{\rm cut}$ is nuanced.  It should be set large enough that truncation effects are not seen for the broadest lines.  But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting. Extreme Doppler shifts could hypothetically send line cores entirely outside the extents of $N_{\rm cut}$ if set too low.  We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA. We coerce all wing cuts to be the same number of pixels, typically 6000 pixels, $\sim30-60~$\AA~ for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line. We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $6000$ pixels per line and assume zeros everywhere else.

@@

Second, another anticipated flaw occurs in the line cores of relatively narrow lines, where the pseudo-Voigt profile becomes a poor approximation of the exact Voigt profile.  The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.

Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen and neutral alkali metal lines.  The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile.  Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.  Narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike. This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.

@@

\subsection{Joint Stellar and Telluric Model}
Figure \ref{blase_flowchart} shows a visual guide to all the steps in  \emph{blas\'e}.  We have arrived at what may be the most intriguing-and-yet-simple of these steps: we simply multiply the rotationally-broadened-and-$RV$-shifted stellar model by the telluric transmission:

>>>>

We retain a relatively small number of pixels $N_{\rm cut}$ adjacent to the line center. Setting this wing cut produces a speedup by a factor of $\frac{N_S}{N_{\mathrm{cut}}}$, which can exceed $100\times$ for wide bandwidth spectra. The choice of $N_{\rm cut}$ is nuanced.  It should be set large enough that truncation effects are not seen for the broadest lines.  But even more, $N_{\rm cut}$ has to be future-proofed for Doppler-shifting. Extreme Doppler shifts could hypothetically send line \edit1{centers} entirely outside the extents of $N_{\rm cut}$ if set too low.  We therefore typically set wingcuts comparable to the buffer size $2 \Delta \lambda_{\mathrm{buffer}}$, even though most weak lines only perceptibly affect $<1\;$\AA. We coerce all wing cuts to be the same number of pixels, typically 6000 pixels, $\sim30-60~$\AA~ for \texttt{PHOENIX}, with the middle pixel being at the line center position, and about 3000 pixels to the red and blue side of the line. We populate a new approximate sparse matrix $\ln{\bm{\hat{F}}}$ with only these $6000$ pixels per line and assume zeros everywhere else.

@@

Second, a conspicuous flaw occurs in the line cores of relatively narrow lines.  The cloned model tends to overestimate the flux at the core and underestimate the flux along the slopes of the lines.  \edit1{The residuals therefore exhibit a \texttt{W}-shape ringing artifact.  The scientific impact of this flaw will depend on the precision demands of the application, but it may not be as bad as it seems for many applications.  Recall that the cloning occurs at native spectral resolution, much greater than typical instrumental resolution.  The \texttt{W}-shape ringing residuals will tend to cancel each other out once the clone model undergoes convolution with an instrumental profile (Section \ref{jointModel}).  We ascribe this flaw to the adoption of approximate pseudo-Voigt profiles rather than exact Voigt profiles.  High-precision applications may wish to employ the exact Voigt profile, which we estimate to cost about 27$\times$ more than this approximate form.}

\edit1{The overprediction of line cores in narrow lines is only partly attributable to the pseudo-Voigt approximation.  Even exact Voigt profiles inadequately represent true atmospheric opacities. Typical stellar and substellar atmosphere models compute opacities for a stratified atmosphere consisting of tens or hundreds of locally isothermal, isobaric atmospheric layers.  Each of those layers may exhibit a local per-line opacity that matches a Voigt profile.  But the cumulative product of many slightly-different Voigt profiles is not exactly equal to any single Voigt profile.  So the \emph{blas\'e} approach of asserting lines as Voigt profiles can be best understood as a phenomenological model-- that lines tend to look Voigt-like, even if they are non-Voigt in detail.  In principle, \texttt{blas\'e} could be configured in a way to learn the requisite perturbations to the Voigt profile in a systematic manner, pooling information among lines.  Such an approach is not yet implemented, and may be an enticing avenue of future research.}

Finally, and most perniciously, a large category of residuals appear near the wings of the deepest and broadest lines---such as hydrogen lines and neutral alkali metal lines.  The true lines exhibit advanced lineshapes, such as non-Lorentzian line wings that are not captured with the overly simplistic Voigt line profile, even an exact one.  Figure \ref{fig_zoom_cloning_performance} highlights super-Lorentzian line wings around a line at 8691~\AA.  \edit1{Nearby} narrow lines devolve into missing line wing opacity, the favored tradeoff when the continuum estimate's poor performance outweighs the pain of a narrow-but-tolerably-small spike. This flaw can be seen where a line initialized at 8692.5~\AA~ and another pair of lines at 8690.0 all melt into line wings.

@@

\subsection{Joint Stellar and Telluric Model}\label{jointModel}
Figure \ref{blase_flowchart} shows a visual guide to all the steps in  \emph{blas\'e}.  We have arrived at what may be the most intriguing-and-yet-simple of these steps: we simply multiply the rotationally-broadened-and-$RV$-shifted stellar model by the telluric transmission:
